---
title: "Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Step 1

Load relevant data (PUMS for persons)
```{r}
library(tidycensus)
library(tidyverse)

person_vars_2018 <- pums_variables %>%
  distinct(year, survey, var_code, 
           var_label, data_type, level) %>%
  filter(level == "person", year == 2018, survey == "acs1")

```

# Step 2

I am going to use the following data: 
  1. Marital Status MAR (Categorical)
  2. Number of hours worked per week WKHP (Continuous - though an argument        can be made for it being categorical since there are only a certain number of hours)
  3. Wages or Salary Income WAGP (Continuous)
  4. Gave birth to a child w/in Past 12 Months FER (Categorical)
  5. Number of Times Married MARHT (Continuous)
  
  Downloading those specific data observations for the state of Nevada from ACS 2018:
  
```{r load person level data, results = 'hide'}
person_data <- get_pums(variables = c("MAR", 
                                      "WKHP", 
                                      "WAGP", 
                                      "FER", 
                                      "MARHT"),
                        state = "NV", 
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  
  select(MAR_label, WKHP, WAGP, FER, MARHT)
```
This results in a dataframe with 28,927 observations on a per person basis for individuals in Nevada in 2018 given 5 different variables. 

# Step 3

 Cleaning up variable names:
  
```{r}
colnames(person_data)<-c("maritalstatus","weeklyhoursworked","salary","children","totalmarriages")
```

# Step 4
 
  Writing as a *csv: 
```{r}
write_csv(person_data,"assignment1.csv")
head(person_data, 10)
```
  
